---
title: "Trabalho 1 - Introducao a Pacotes Estatisticos"
output: html_notebook
---

##Leitura do arquivo e renomeação das variáveis

####Ler o arquivo .csv e carregar os dados para a variavel 'data'. 

```{r}
data <- read.csv2("data.csv", dec=".", sep=",", encoding = 'UTF-8');
```

Renomear as colunas para nomes mais sucintos.

```{r}
names(data)[names(data) == "Data.de.nascimento."] <- "nascimento";
names(data)[names(data) == "Gênero"] <- "genero";
names(data)[names(data) == "Qual.a.sua.escolaridade."] <- "escolaridade";
names(data)[names(data) == "Qual.a.sua.altura."] <- "altura";
names(data)[names(data) == "Qual.o.seu.peso..em.kg.."] <- "peso";
names(data)[names(data) == "Quantas.horas.em.média.você.costuma.dormir.por.noite."] <- "horas_sono";
names(data)[names(data) == "Quantas.vezes.na.semana.em.média.você.pratica.exercícios.físicos."] <- "exercicios_fisicos";
names(data)[names(data) == "Com.que.frequência.você.consome.alimentos.com.grande.quantidade.de.açúcar."] <- "acucar";
names(data)[names(data) == "Com.que.frequência.você.consome.alimentos.de.origem.animal."] <- "origem_animal";
names(data)[names(data) == "Você.é.estudante."] <- "estudante";
names(data)[names(data) == "Qual.turno.você.estuda."] <- "turno";
names(data)[names(data) == "Quantas.horas.por.semana..em.média..você.estuda.em.casa."] <- "horas_estudo";
names(data)[names(data) == "Como.você.avalia.o.seu.desempenho.escolar.acadêmico."] <- "desempenho_estudo";
names(data)[names(data) == "Você.está.atualmente.empregado.a.."] <- "empregado";
names(data)[names(data) == "Qual.o.tipo.de.trabalhador.você.é."] <- "tipo_trabalhador";
names(data)[names(data) == "Como.você.considera.o.seu.rendimento.na.sua.atual.ocupação."] <- "desempenho_trabalho";
names(data)[names(data) == "Com.que.frequência.você.se.sente.cansado.a.."] <- "cansado";
names(data)[names(data) == "Quantas.vezes.no.último.ano.você.ficou.doente."] <- "vezes_doente";
names(data)[names(data) == "Nas.vezes.que.ficou.doente..foi.necessário.ir.a.um.médico."] <- "foi_medico";
names(data)[names(data) == "Depois.de.ir.ao.médico..você.precisou.tomar.alguma.medicação."] <- "medicacao";
names(data)[names(data) == "Você.já.doou.sangue."] <- "doador";
names(data)[names(data) == "Você.já.se.sentiu.mal.ao.doar.sangue."] <- "doador_sentiu_mal";
names(data)[names(data) == "Mais.alguém.da.sua.família.já.se.sentiu.mal.ao.doar.sangue."] <- "doador_familia";

```



Salvar arquivo.

```{r}
write.csv(data, file="data_renamed_columns.csv");
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

